---
title: "BORIS_analyses"
output: 
  html_document: 
    theme: cerulean
    number_sections: yes
date: "2023-11-27"
---

```{r include=FALSE}
# install.packages("tidyverse", dependencies = T)
# install.packages("ggthemes")
library(tidyverse)
library(ggthemes)
library(colorspace)

#---- DATA FROM DATASHEETS ----

#export the list of all trials r1, r2 and r3 for the meeting early December
mtg_trial_ID <- read_tsv("./Full_trial_list.tsv") |> 
  filter(Replicates %in% c("r1","r2","r3")) |> 
  filter(Camera_nb == "Camera2") |> 
  filter(Data_analysis != "N") |> 
  pull(ID)

#import the 3 files for Small, Medium and Large size classes from datasheets
datasheet_files <- read_tsv(c("Treadmill_Fish_Contact - Trial_Observations_cohort2_small_6-10cm.tsv",
                            "Treadmill_Fish_Contact - Trial_Observations_cohort2_medium_11-15cm.tsv",
                            "Treadmill_Fish_Contact - Trial_Observations_cohort2_large_16-20cm.tsv"))
colnames(datasheet_files)
#keep only metadata for BORIS analysis
metadata <- datasheet_files |> 
  select(c(Trial_ID_cam2,Trial_ID_cam4,Size_class,Replicate,Treatment,Dial_phase,Water_Temp_Tread,
           DO_Tread_mgL,Water_Temp_JC,DO_JC_mgL,Lux,Video_File_Name)) |> 
  distinct() #remove duplicates in rows
#double check
sum(metadata$Size_class == "Small")
sum(metadata$Size_class == "Medium")
sum(metadata$Size_class == "Large")

#filter by mtg_trial_ID (keep only r1, r2 and r3 trials)
boris_metadata <- metadata |> 
  filter(Trial_ID_cam2 %in% mtg_trial_ID)
```

# Importing data

Double check the number of videos analyzed per size class.

Considering only r1, r2 and r3 videos, we should have for Small, Medium and Large respectively 25/25/24.

```{r echo=TRUE}
sum(boris_metadata$Size_class == "Small")
sum(boris_metadata$Size_class == "Medium")
sum(boris_metadata$Size_class == "Large")
```

```{r include=FALSE}
#---- DATA FROM BORIS ----

#import the list of the .tsv files exported from BORIS (only from video analyzed)
HD_file_list <- list.files(".", pattern = "cam2", recursive = T)
#filter by mtg_trial_ID (keep only r1, r2 and r3 trials)
HD_file_list = grep(paste(mtg_trial_ID, collapse="|"), HD_file_list, value = TRUE)

#change column names using a vector while creating a large tibble to avoid errors
col_names <- c("Observation_id","Observation_date","Description","Observation_type",
               "Source","Total_duration","Media_duration_s","FPS_frame_s","Date",
               "Trial_Letter","Camera_#","Subject","Observation_duration_by_subject_by_observation",
               "Behavior","Behavioral_category","Behavior_type","Start_s","Stop_s","Duration_s",
               "Media_file_name","Image_index_start","Image_index_stop","Image_file_path_start",
               "Image_file_path_stop","Comment_start","Comment_stop")

#create a tibble with all the data 
boris_files <- read_tsv(HD_file_list, skip = 1, col_names = col_names, col_types = "cTcccdddccccdcccdddcddcccc") %>%
  mutate(Date = as.Date(Date, format = "%Y%m%d")) |> 
  select(-c(Observation_date,Description,Observation_type,Image_index_start,Image_index_stop,Image_file_path_start,Image_file_path_stop))
#double check the tibble and column types
str(boris_files)
#if any errors
problems(boris_files)

#---- MERGE BOTH TIBBLES ----

boris <- merge(boris_files, boris_metadata, by.x = "Observation_id", by.y = "Trial_ID_cam2")
```


Visualization of the whole dataset:

```{r echo=FALSE}
glimpse(boris)
```


Double check consistency between video file name manually entered (column called "Video_File_Name") and video file path (column called "Media_file_name").

```{r include=FALSE}
#double check consistency between video and media file names columns (because entered manually 2 different times)
boris$Media_file_name
boris$Video_File_Name
#loop to check for each row if the pattern Video_File_Name is well found in the Media_file_name path
tmp = map2_lgl(
  boris$Video_File_Name,
  boris$Media_file_name,
  function(pattern, x) grepl(pattern, x, ignore.case = T) #looking for the pattern (video file name) in each x (media file name)
)
```
=> number of errors:
```{r echo=FALSE}
sum(!tmp)

#export tibble
#write_tsv(boris, "BORIS_large_dataset_r1r2r3.tsv")
```


# What type of velocity treatments make them fatigued faster?


**TO DO: remove rows where fish were resting then probed**


```{r include=FALSE}
#---- FIGURES ----

#Prep data

#data
tmp <- boris |> 
  filter(Behavior == "resting") |> 
  group_by(Size_class, Dial_phase, Treatment)
#check column types
typeof(tmp$Duration_s)
tmp$Treatment <- factor(tmp$Treatment, levels = c("0.2A_0.8S","0.2A_2S","0.4A_0.8S","0.4A_2S"))
tmp$Size_class <- factor(tmp$Size_class, levels = c("Small","Medium","Large"))
tmp$Dial_phase <- factor(tmp$Dial_phase, levels = c("Day","Night"))

#remove rows where fish were resting then were probed
#TO DO!!!!!
```

## Number of fish resting per treatment

```{r echo=FALSE, fig.height=4, fig.width=5}
ggplot(tmp, aes(x = Treatment, fill = Treatment)) +
  geom_bar() +
  labs(x = "Treatment", y = "Resting fish counts") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_discrete_qualitative(palette = "Set 3") + #https://arxiv.org/pdf/1903.06490.pdf
  theme_classic()
#ggsave("figures/resting_total_nb.pdf", width = 5, height = 4)
```

Results below:

```{r echo=FALSE}
#double check with values: nb of fish resting per treatment
boris |> 
  filter(Behavior == "resting") |> 
  group_by(Treatment) |> 
  count() #nb of rows = nb of fish resting
```


## Number of fish resting per treatment per size class

```{r echo=FALSE, fig.height=5, fig.width=10}
#figure 2
ggplot(tmp, aes(x = Treatment, fill = Treatment)) +
  geom_bar() +
  facet_grid(cols = vars(Size_class)) +
  labs(x = "Treatment", y = "Resting fish counts") +
  scale_y_continuous(limits = c(0, 60), expand = c(0, 0)) +
  scale_fill_discrete_qualitative(palette = "Set 3") + #https://arxiv.org/pdf/1903.06490.pdf
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
#ggsave("figures/resting_nb_per_sizeclass.pdf", width = 10, height = 5)
```

Results below:

```{r echo=FALSE}
#double check with values: nb of fish resting per treatment per size class
boris |> 
  filter(Behavior == "resting") |> 
  group_by(Size_class, Treatment) |> 
  count() #nb of rows = nb of fish resting
```


## Number of fish resting per treatment per size class per dial phase

```{r echo=FALSE, fig.height=7, fig.width=7}
#figure 3
ggplot(tmp, aes(x = Treatment, fill = Treatment)) +
  geom_bar() +
  facet_grid(vars(Size_class), vars(Dial_phase)) +
  labs(x = "Treatment", y = "Resting fish counts") +
  scale_y_continuous(limits = c(0, 40), expand = c(0, 0)) +
  scale_fill_discrete_qualitative(palette = "Set 3") + #https://arxiv.org/pdf/1903.06490.pdf
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
#ggsave("figures/resting_nb_per_sizeclass_and_dial_phase.pdf", width = 7, height = 7)
```

Results below:

```{r echo=FALSE}
#double check with values: nb of fish resting per treatment per size class per dial phase
boris |> 
  filter(Behavior == "resting") |> 
  group_by(Size_class, Dial_phase, Treatment) |> 
  count() #nb of rows = nb of fish resting
```


## Duration of fish resting per treatment

```{r echo=FALSE, fig.height=5, fig.width=6}
#figure 4
ggplot(tmp, aes(x = Treatment, y = Duration_s, fill = Treatment)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.21) +
  labs(x = "Treatment", y = "Duration of fish resting (s)") +
  scale_y_continuous(limits = c(0,470), expand = c(0, 0)) +
  scale_fill_discrete_qualitative(palette = "Set 3") + #https://arxiv.org/pdf/1903.06490.pdf
  theme_classic()
#ggsave("figures/resting_total_duration.pdf", width = 6, height = 5)
```

Results below:

```{r echo=FALSE}
#double check with values: mean duration of fish resting per treatment
boris |> 
  filter(Behavior == "resting") |> 
  group_by(Treatment) |> 
  summarise(duration_µ = mean(Duration_s),duration_sd = sd(Duration_s))
```


## Duration of fish resting per treatment per size class

```{r echo=FALSE, fig.height=5, fig.width=10}
#figure 5
ggplot(tmp, aes(x = Treatment, y = Duration_s, fill = Treatment)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.21) +
  facet_grid(cols = vars(Size_class)) +
  labs(x = "Treatment", y = "Duration of fish resting (s)") +
  scale_y_continuous(limits = c(0,470), expand = c(0, 0)) +
  scale_fill_discrete_qualitative(palette = "Set 3") + #https://arxiv.org/pdf/1903.06490.pdf
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
#ggsave("figures/resting_duration_per_sizeclass.pdf", width = 10, height = 5)
```

Results below:

```{r echo=FALSE}
#double check with values: mean duration of fish resting per treatment per size class
boris |> 
  filter(Behavior == "resting") |> 
  group_by(Size_class, Treatment) |> 
  summarise(duration_µ = mean(Duration_s),duration_sd = sd(Duration_s))
```


## Duration of fish resting per treatment per size class per dial phase

```{r echo=FALSE, fig.height=7, fig.width=7}
#figure 6
ggplot(tmp, aes(x = Treatment, y = Duration_s, fill = Treatment)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.21) +
  facet_grid(vars(Size_class), vars(Dial_phase)) +
  labs(x = "Treatment", y = "Duration of fish resting (s)") +
  scale_y_continuous(limits = c(0,470), expand = c(0, 0)) +
  scale_fill_discrete_qualitative(palette = "Set 3") + #https://arxiv.org/pdf/1903.06490.pdf
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
#ggsave("figures/resting_duration_per_sizeclass_and_dial_phase.pdf", width = 7, height = 7)
```

Results below:

```{r echo=FALSE}
#double check with values: mean duration of fish resting per treatment per size class per dial phase
boris |> 
  filter(Behavior == "resting") |> 
  group_by(Size_class, Dial_phase, Treatment) |> 
  summarise(duration_µ = mean(Duration_s),duration_sd = sd(Duration_s))
```

