---
title: "Boris Analysis - 2023 dataset"
author: "Anna Steel"
date: "11/28/2023"
output: html_document
---

```{r setup}
library(tidyverse)
library(lme4)
library(patchwork)
library(chron)
library(MuMIn)

```
  
  
## Read in Dataset  
This code was started with only r1,r2,r3 from Camera 2 (Nov 28, 2023)
```{r read data, echo=F}
dat.raw = read.csv("/Users/Anna/Documents/ResearchGit/Treadmill/BORIS/rawdata/BORIS_large_dataset_r1r2r3_nameedit.csv")

```

```{r reduce columns and melt into shape}

dat.pre = dat.raw %>% 
  select(!c(Source, Media_duration_s, FPS_frame.s, Subject, 
            Observation_duration.by.subject.by.observation,
            Behavior_type, Media_file_name, Trial_ID_cam4, 
            Comment_start, Comment_stop)) %>%
  mutate(BehaviorID=paste(Video_File_Name, Start_s))

 
dat.contact = dat.pre %>%
  filter(Behavioral_category!="resting") %>%
  pivot_wider(names_from = Behavioral_category, values_from = Behavior) %>%
  mutate(resting=NA)

dat.rest = dat.pre %>%
  filter(Behavioral_category=="resting") %>%
  rename(resting = Behavioral_category) %>%
  select(-Behavior) %>%
  mutate(bodypart=NA, contacts=NA, orientation=NA)

dat.rest = dat.rest[,names(dat.contact)] # reorganize to match

dat = rbind(dat.rest, dat.contact)

dat = arrange(dat, Date, Diel_phase, Trial_Letter, Start_s)

dat$Treatment = factor(dat$Treatment, levels=c("0A_0S", "0.2A_0.8S", "0.2A_2S", "0.4A_0.8S", "0.4A_2S"))
```

#### A Few Interesting Things:
- There are a total of 1967 records in these first three replicated trials that do not have orientation but DO have bodypart and contact
    - For example, Video 2023Aug14_1220PM_C doesn't have any orientation data collected


## Exploring the contacts

### Total contact rate
```{r contact rate}
sumdat = dat %>%
  mutate(treatID = paste(Size_class, Diel_phase, Treatment, sep="_")) %>%
  group_by(treatID, Size_class, Diel_phase, Treatment, Replicate) %>%
  summarize(ncontact = n())

contact_rate.v1 = ggplot(sumdat, aes(x=factor(Diel_phase), y=ncontact, fill=Treatment))+
  geom_boxplot() + 
  scale_fill_manual(values=c("grey80","coral2","skyblue","brown3","steelblue3")) +
  ylab("N contacts per trial") + xlab("Diel Phase") + 
  facet_grid(Size_class~.) +
  theme_bw() #+ theme(axis.text.x = element_text(angle=45, vjust=.85, hjust=1))

contact_rate.v1 # + geom_vline(xintercept = 1.5, lty="dashed")


contact_rate.v2 = ggplot(sumdat, aes(x=factor(Treatment), y=ncontact, fill=Diel_phase))+
  geom_boxplot() + 
  #scale_fill_manual(values=c("grey80","coral2","skyblue","brown3","steelblue3")) +
  ylab("N contacts per trial") + xlab("Diel Phase") + 
  facet_grid(Size_class~.) +
  theme_bw() #+ theme(axis.text.x = element_text(angle=45, vjust=.85, hjust=1))

contact_rate.v2  + geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5), lty="dashed", alpha=.5)

```


### Body/Head/Tail
```{r body-head-tail contacts}
sumdat.body = dat %>%
  filter(!is.na(bodypart)) %>%
  group_by(Video_File_Name, Date, Diel_phase, Trial_Letter, Camera, 
           Size_class, Replicate, Treatment, Water_Temp_Tread, bodypart) %>%
  summarize(Freq=n()) %>%
  data.frame()



body.small = ggplot(filter(sumdat.body, Size_class=="Small"), 
                    aes(x=Treatment, y=Freq, fill=Treatment))+
  geom_boxplot() + 
  scale_fill_manual(values=c("grey80","coral2","skyblue","brown3","steelblue3"), guide=F) +
  ylab("N contacts per trial") + xlab("Flow Treatment")+ ggtitle("Contact Type, Small Size Class (6-10cm TL)") + 
  facet_grid(bodypart~Diel_phase) +
  theme_bw() + theme(axis.text.x = element_text(angle=45, vjust=.85, hjust=1))
  

body.medium = ggplot(filter(sumdat.body, Size_class=="Medium"), 
                    aes(x=Treatment, y=Freq, fill=Treatment))+
  geom_boxplot() + 
  scale_fill_manual(values=c("grey80","coral2","skyblue","brown3","steelblue3"), guide=F) +
  ylab("N contacts per trial") + xlab("Flow Treatment")+ ggtitle("Contact Type, Medium Size Class (11-15cm TL)") + 
  facet_grid(bodypart~Diel_phase) +
  theme_bw() + theme(axis.text.x = element_text(angle=45, vjust=.85, hjust=1))


body.large = ggplot(filter(sumdat.body, Size_class=="Large"), 
                    aes(x=Treatment, y=Freq, fill=Treatment))+
  geom_boxplot() + 
  scale_fill_manual(values=c("grey80","coral2","skyblue","brown3","steelblue3"), guide=F) +
  xlab("Flow Treatment")+ 
  ylab("N contacts per trial") + ggtitle("Contact Type, Large Size Class (16-19cm TL)") + 
  facet_grid(bodypart~Diel_phase) + 
  theme_bw() + theme(axis.text.x = element_text(angle=45, vjust=.85, hjust=1))


body.small + body.medium + body.large + plot_layout(ncol=3)



## rearrange thr groupings to highglight bodypart

body.small.b = ggplot(filter(sumdat.body, Size_class=="Small"), 
                    aes(x=bodypart, y=Freq, fill=bodypart))+
  geom_boxplot() + 
  scale_fill_manual(values=c("grey80","coral2","steelblue3"), guide=F) +
  xlab("Contact Type")+ 
  ylab("N contacts per trial") + ggtitle("Contact Type, Small Size Class (6-10cm TL)") + 
  facet_grid(Treatment~Diel_phase) +
  theme_bw() + theme(axis.text.x = element_text(angle=45, vjust=.85, hjust=1))
  

body.medium.b = ggplot(filter(sumdat.body, Size_class=="Medium"), 
                    aes(x=bodypart, y=Freq, fill=bodypart))+
  geom_boxplot() + 
  scale_fill_manual(values=c("grey80","coral2","steelblue3"), guide=F) +
  xlab("Contact Type")+ 
  ylab("N contacts per trial") + ggtitle("Contact Type, Medium Size Class (11-15cm TL)") + 
  facet_grid(Treatment~Diel_phase) +
  theme_bw() + theme(axis.text.x = element_text(angle=45, vjust=.85, hjust=1))


body.large.b = ggplot(filter(sumdat.body, Size_class=="Large"), 
                    aes(x=bodypart, y=Freq, fill=bodypart))+
  geom_boxplot() + 
  scale_fill_manual(values=c("grey80","coral2","steelblue3"), guide=F) +
  xlab("Contact Type")+ 
  ylab("N contacts per trial") + ggtitle("Contact Type, Large Size Class (16-19cm TL)") + 
  facet_grid(Treatment~Diel_phase) + 
  theme_bw() + theme(axis.text.x = element_text(angle=45, vjust=.85, hjust=1))


body.small.b + body.medium.b + body.large.b + plot_layout(ncol=3)

```

### Rheotaxis
```{r rheotaxis}

```


### Body part x Rheotaxis
```{r contact type interaction}
```


### Relationships between contact rates/types and controlled variables
- temperature in treadmill, mean size per trial, order of trial, temperature difference, more?
```{r experimental set-up}
```



